---
title: "Covid regression"
author: ""
date: "8/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Regression problem

- We will run regression and other related models for Covid-19 data

## Libiraries

```{r}
library(tidyverse)
library(caret)
library(glmnet)
```


## Check data

- check the distribution of the output (deaths_per1000)
- need conversion?

```{r}

data_covid <- read_csv("data/covid-data.csv.gz")
head(data_covid)
```

## Check the distribution of output

```{r}
ggplot(data_covid, aes(x = deaths_per1000)) + geom_density()
```

```{r}
ggplot(data_covid, aes(x = deaths_per1000 +.01)) + geom_density() + scale_x_log10()
```


## Decide the variable to include as input

- There are 47 variables what are possible predictors?
  - trump: pct_report, votes, total_votes, pct, lead, 
  - demography: TotalPop, Men, Women, Hispanic, White, Black, Native, Asian, Pacific, VotingAgeCitizen, Income, IncomePerCap, Poverty, ChildPoverty, Professional, Service, Office, Construction, Production, Drive, Carpool, Transit, Walk, OtherTransp, WorkAtHome, MeanCommute, Employed, PrivateWork, PublicWork, SelfEmployed, FamilyWork, Unemployment
  - deaths_per1000

```{r}
data_covid_use <- data_covid %>%
  mutate(deaths_per1000_log = log(deaths_per1000 + 0.01)) %>%
  select(deaths_per1000_log, Black, Hispanic, IncomePerCap,
         Poverty, Unemployment, pct) 
  
```

## Train-test split

```{r}
train_id <- createDataPartition(data_covid_use$deaths_per1000_log, list = F, p = .7)
data_train <- data_covid_use %>% slice(train_id)
data_test <- data_covid_use %>% slice(-train_id)

```

## Preprocess

```{r}
prep <- data_train %>% select(-deaths_per1000_log) %>%
  preProcess(method = c("center", "scale"))

data_train_preped <- predict(prep, data_train)
data_test_preped <- predict(prep, data_test)

```

## Linear regression

```{r}
model_lm <- lm(deaths_per1000_log ~ ., data = data_train_preped)
data_train_preped %>%
  mutate(pred_lm = predict(model_lm)) %>%
  ggplot(aes(x = pred_lm, y = deaths_per1000_log)) + geom_point()

summary(model_lm)

```
```{r}
rmse_lm_trian <- (predict(model_lm) - data_train_preped$deaths_per1000_log)^2 %>% mean()
rmse_lm_test <- (predict(model_lm, newdata = data_test_preped) - 
                   data_test_preped$deaths_per1000_log)^2 %>% mean()

```
## KNN-regression

```{r}
ctrl <- trainControl(method = "repeatedcv",
                     number = 5,
                     repeats = 3)
model_knn <- train(deaths_per1000_log ~ ., data = data_train_preped, 
   method = "knn", trControl = ctrl)

model_knn
pred_train_knn <- predict(model_knn)
pred_test_knn <- predict(model_knn, newdata = data_test_preped)
rmse_trian_knn <- (pred_train_knn - data_train_preped$deaths_per1000_log)^2 %>% mean()
rmse_test_knn <- (pred_test_knn - 
                   data_test_preped$deaths_per1000_log)^2 %>% mean()

```

## Ridge and LASSO

```{r}
    boston_ridge_cv <- cv.glmnet(Boston_train_X, 
                                 Boston_train_Y, 
                                 alpha = 0,
                                 type.measure = "mse", 
                                 family = "gaussian")
    ```


## Ridge regression outcome (MSE)

```{r eval = F}
plot(boston_ridge_cv)
data_train_x <- data_train_preped %>% select(!deaths_per1000_log) %>% as.matrix()
data_test_x <- data_test_preped %>% select(!deaths_per1000_log) %>% as.matrix()
model_ridge <- cv.glmnet(data_train_x, data_train_preped$deaths_per1000_log, data = data_train_preped, 
                             alpha = 0,
                                 type.measure = "mse", 
                                 family = "gaussian")

plot(model_ridge)
model_ridge_cv
pred_train_ridge <- predict(model_ridge, data_train_x)
pred_test_ridge <- predict(model_ridge, newx = data_test_x)
rmse_trian_ridge <- (pred_train_ridge - data_train_preped$deaths_per1000_log)^2 %>% mean()
rmse_test_ridge <- (pred_test_ridge - 
                   data_test_preped$deaths_per1000_log)^2 %>% mean()
plot(model_ridge$glmnet.fit, xvar = "lambda")

```

```{r}
model_lasso <- cv.glmnet(data_train_x, data_train_preped$deaths_per1000_log, data = data_train_preped, 
                             alpha = 1,
                                 type.measure = "mse", 
                                 family = "gaussian")

plot(model_lasso)
model_lasso
pred_train_lasso <- predict(model_lasso, data_train_x)
pred_test_lasso <- predict(model_lasso, newx = data_test_x)
rmse_trian_lasso <- (pred_train_lasso - data_train_preped$deaths_per1000_log)^2 %>% mean()
rmse_test_lasso <- (pred_test_lasso - 
                   data_test_preped$deaths_per1000_log)^2 %>% mean()
plot(model_lasso$glmnet.fit, xvar = "lambda")
coef(model_lasso)
```


## Final Verdict